---
title: "Constructing your own feature detector"
author: "Kohleth Chia"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Constructing your own feature detector}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warnings=FALSE
)
```

```{r,include=FALSE}
library(dplyr) 
```
Although the package ships with its own set of feature detectors. It is actually not that difficult to incorporate your own.
All the shipped feature detectors are functions whose name starts with `.dt`. See `?definitions`. 

## A naive example
In principle, feature detectors are just functions which take a data.frame and return another with columns `curve_id`, `Cl:xxx`, and `fname`, where `Cl:xxx` is the clause component of the feature definition, and `fname` is the machine-friendly feature name. So, for example,
```{r}
library(gaitFeature)
```

```{r}
myRandomDetector=function(x){
  detected=sample(0:1,nrow(x),TRUE)
  data_frame(curve_id=x$curve_id,
             "Cl:random"=as.logical(detected),
             Random=detected)
}

myRandomDetector(kinematics)

```

And you can use the `detectAll` function with it.
```{r}
detectAll(kinematics,fDict=data.frame(fn="myRandomDetector"))%>%
  select(-(`0`:`100`))
```

## More realistic feature detector

However, a useful feature detector will probably be more sophisticated than that. 
And this package provides some useful tool to construct these complicated features. 
But these tools assume a particular way of constructing feature detectors. 
In particular, a feature detector function should consist of 3 components:

1. A filter to a specific joint and plane.
2. One or more clauses.
3. A finish function.

## 1. Filter
A feature should only apply to a particular joint and plane, so we recommend that you explicitly incorporate a filter in the definition.

## 2. Clauses
A feature definition should consists of one or more *clauses*. 
The most commonly used clause is what we called a *univariate statistics clauses*.

And it works as followed:

1. From a curve select a time domain
2. Compute a summary statistics on the selected curve segment.
3. Compute the same statistics on each of the reference curve (for typically developed children).
4. If the statistics computed in 2. is extreme* compared to the set of statistics computed in 3., the clause is passed.

*The definition of extreme is user-definable. For example, it can be > mean + 2 standard deviation.

Step 3 above calls for a set of reference curve. This packages ships with such a set, and can be called by the `getRC()` function. See also `?.rc`. 

This clause is implemented in the function `unistat`. See `?unistat`.

For example ``unistat(x,`60`:`100`, .f=mean, .dir=">", .k=2)`` reads the mean angle at t in [60,100] should be greater than mean + 2SD of the same mean copmuted from the typical reference data.

- `.f` can be any function which returns a single number statistics. See `?.tp` for commonly used functions that ships with this package.
- The possible directions `.dir` are: `>(=)`, `<(=)`, `%w%`, and `%o%`, where the last two stands for 'within' and 'outside'.
- An absolute threshold can be specified by using the `.c=xxx` argument instead of `.k=xx`. 


## 3. Finish with `finish`
A feature detector function should finish off with the `finish` function, which does several house-keeping functions, such as checking all clauses are passed before the feature is deemed present, and setting various attributes appropriately.


## Example
If we look at the detector for increased ankle rotation.
```{r}
.dtAIR
```

We will see that,

1. it is a feature which applies to the Ankle on the transverse plane.
2. There is one clause, which reads the mean angle at t in [60,100] should be greater than mean + 2SD (default) of the same mean copmuted from the typical reference data.
3. The machine-friendly name for this feature is 'AIR', and the human interpretable name is 'Ankle Internal Rotation'.

## Multiple clauses
Multiple clauses can be chained together by the pipe `%>%` operator.
For example,
```{r}
.dtDecPelTiltIncROM
```
has 2 clauses.

## Other clauses
In addition to `unistat`, we provide two more clause functions: `corr` and `custom`.
These clauses become useful when the feature is difficult to define through `unistat`.


### Correlation clause `corr`
A correlation clause works as follow:

1. From a curve select a time domain.
2. Correlate that curve segment with a set of pre-supplied target curves.
3. If any of the correlation meets the target threshold the clause is passed.

Therefore, in order to use this clause, a set of target curves needs to be supplied (using the `corr(..., TC=x,...)`).

This package ships with 5 sets of target curves, which can be assessed by the `.tc()` function.
They are: `.tc()$ett`, `.tc()$uni`, `.tc()$s2r`, `.tc()$revrom`, and `.tc()$dblbump`, which are target curves for
'External Tibial Torsion', 'Uni Bump', 'Short 2nd rocker', 'Reverse ROM', and 'Double Bump'.

Since all these curves have domain 0:100, the same domain should be specified in the `corr` function.

Straightly speaking, we could use the `unistat` clause to achieve the same result. That is,

`corr(...,TC=x,">=",0.8)` is equivalent to

``unistat(`0`:`100`,.f=function(x)max(apply(t(.tc()$dblbump),2,cor,x)),.dir=">=",.c=0.8)``

(the `apply` function is only needed because we have a set of `TC`. Otherwise a simple `cor` will suffice.)

However, the `corr` clause is just more literal and thus easier to interpret.

### Custom clause `custom`
The custom clause function `custom` allows one to supply a custom condition through as a string `custom(...,cond)`.
It is particularly useful for clauses that require data other than the kinematics, such as timing related clauses. For example,
```{r}
.dtDelayPkKnFx
```
has a clause which says the peak angle after `t=60` must occur after `t=75`.

Again, straightly speaking, this clause can be expressed as a univariate clause. That is,

``custom(`60`:`100`, cond="t[which.max(angle)]>75")`` 

is equivalent to 

``unistat(`60`:`100`,.f=function(x)which.max(x),.dir=">",.c=75-60+1)``.
 
However, the `unistat` version relies on the assumption that `x` in `.f` is sorted inn `t`, whereas `custom` is explicitly extracting `t`.


## Working with `detectAll`
Suppose you have created your own detector using the steps outlined above. For illustration we simply copy the `.dtAIR` detector. 
```{r}
myDetector=function(df){
  df%>%
    filter(joint=="Ankle"&plane=="tra")%>%
    unistat(`0`:`60`,.f=mean,.dir=">")%>%
    finish(featurename = "My Detector. Should be same as .dtAIR",fname = "MyDetector")
}
```

You can ask `detectAll` to run it by constructing the corresponding feature dictionary.
```{r}
myDict=makeFDict("myDetector")
kinematics%>%
  filter(joint=="Ankle",plane=="tra")%>%
  detectAll(fDict=myDict)%>%
  select(-(`0`:`100`))
```
You can also add it to an existing (e.g. the default) dictionary.
```{r}
defaultDict=makeFDict(".dtAIR")
newDict=defaultDict%>%
  rbind(myDict)

kinematics%>%
  filter(joint=="Ankle",plane=="tra")%>%
  detectAll(fDict=newDict)%>%
  select(-(`0`:`100`))
```
