---
title: "QuickStart"
author: "Kohleth Chia"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{QuickStart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warnings=FALSE
)
```

## Preeliminaries
This is an R package so I assume you are already familiar with R.

Throughout this guide we will be using the dataset `kinematics` which comes with this package.

## Installation
Will finalize this once this repository has a permanent resting place.


## Quick Start
The functions in this package assume the kinematics data is already loaded in R, in the **wide** format. Specifically, the kinematics `data.frame` should contain at least columns `curve_id`, `joint`, `plane`, `0,1,2,...,100`. Each row in the dataframe represent a kinematics curve on a particular joint and plane, identified by field `curve_id`, with data at 101 time points `0,1,2,...,100`.

The package ships with an example data `kinematics`.
```{r}
library(gaitFeature)
library(dplyr) 
library(tidyr)
kinematics
```

The simplest use is to run the `detectAll` function which runs all the off-the-shelf feature detectors on the supplied dataframe.
```{r}
detected=detectAll(kinematics)
detected%>%select(-(`0`:`100`))
```

`detectAll` returns the same data.frame as what is supplied, with extra columns -- one for each feature. If the feature column is `NA`, it means that feature is not applicable to that curve (e.g. Ankle Internal Rotation (AIR) not relevant for the Pelvis curve.) If the field is `1` then that feature is detected, and `0` indicates the contrary.

The list of feature detectors that are called by `detectAll` is specified through the `fDict` argument. This argument should be a data.frame with column `fn` which list the names of the detectors to be called. Furthermore, it can have columns `joint` and `plane` which indicates which joint and plane a particular detector is related to. If `fDict=NULL` (default), it is generated on the spot by the `makeFDict` function.

```{r}
makeFDict()
```

If you are only interested at features that are a particular joint or plane, you can specify that in the `detectAll` call. But make sure the joint and plane value appears in the `fDict` data.frame.

```{r}
detectAll(kinematics, joint="Hip", plane="sag")%>%
  select(-(`0`:`100`))
```
Note that all of `kinematics` will still be returned. So the above result is different to 
```{r}
kinematics%>%
  filter(joint=="Hip",plane=="sag")%>%
  detectAll(joint="Hip",plane="sag")%>%
  select(-(`0`:`100`))
```

A (more) detailed explanation of why each feature is detected or not can be returned by setting `detectAll(...,comp=TRUE)` and inspecting the attributes.
```{r}
z=detectAll(kinematics,joint="Hip",plane="sag",comp=TRUE)
head(attr(z,"comp"),1)
```

`attr(.,"comp")` returns one dataframe for each feature. The columns that begin with `Cl:` indicates it is a clause (component) of the feature. All clauses must be `TRUE` for a feature to be detected.

You can also call the individual detector directly.
```{r}
.dtDecHipFxIC(kinematics)
```


## Resampling
All functions in this package assume the kinematics curves are sampled at 101 time points.
If this is not the case, there is a sampling function `?resamp101` which fits a periodic cubic spline to your existing curves, and then resample the 101 times points from it.

```{r}
## x is a coarser version of kinematics
x=kinematics%>%
  select(curve_id,`0`,`10`,`20`,`30`,`40`,`50`,`60`,`70`,`80`,`90`,`100`)

y=resamp101(x,`0`:`100`,plot=T)
```

The `plot=T` argument plots the interpolated lines together with the observed points. 
It is recommended that you leave it as `FALSE` (default) if you are resampling many curves, because the plot will just be cluttered.
